# Methods/Data
```{r, echo = FALSE}
source("scripts/methods_tables.R")
```

## Methods

### Phasing

-axiom
-omni
-coreExome

### SelectionTools



Selection statistics that are generated:

Window based
- Tajima's D (vcftools) [@Tajima1989]
- Fay and Wu's H [@fay2000hitchhiking] (variscan)
- Fu and Li's D [@Fu1993] (variscan)
- Fu and Li's F [@Fu1993] (variscan)
- Eta (variscan) total (minimum number) number of mutations ($\eta$) 
- Eta E (variscan) number of external mutations ie number of derived singletons
- Pi (variscan) nucleotide diversity ($\pi$)
- S (variscan) number of segregating sites
- mean F~ST~ (vcftools)

Haplotype based
- sL0
- sL1
- unstandardised nSL
- standardised nSL
- ihh0
- ihh1
- unstandardised iHS
- standardised iHS
- iHH population 1
- iHH population 2
- unstandardised XP-EHH
- standardised XP-EHH

Frequency
- Minor allele frequency
- Derived allele frequency

Gene
- Ka/Ks




#### Selection database creation


### GWAS

Uk Biobank GWAS


####

#### Heritability Analysis

10000 controls were randomly selected. The same 10000 samples were then combined with each different gout criteria and heritiability estimates were obtained by first creating the genetic relationship matrix and then the variance of gout 
```{r, engine='bash', eval=FALSE}
# load fam into R
# sample(10000, fam[AFF == controls]) -> controls.txt
# awk '{if($6 == 2){print $1"\t"$2}}' < condition.fam | cat - controls.txt > ids_keep.txt

# Subset samples
for i in $(seq 1 22)
do 
  plink2 --bed /Volumes/BiocArchive/archive/merrimanlab/central_datasets/ukbiobank/plink/chr${i}impv1.bed \
            --bim /Volumes/BiocArchive/archive/merrimanlab/central_datasets/ukbiobank/plink/chr${i}impv1.bim_1kg_marker \
            --fam /Volumes/BiocArchive/archive/merrimanlab/central_datasets/ukbiobank/plink/chrallimpv1.fam_allgout_allcontrols \
            --hwe 0.000001 \
            --maf 0.01 \
            --keep ult_ids_keep.txt \
            --make-bed \
            --out gout_chr${i} \
done
```

Calculate the genetic relationship matrix
```{r, engine = 'bash', eval = FALSE}
for i in $(seq 1 22)
do 
  gcta1.26.0 \
    --bfile gout_chr${i} \
    --chr ${i} \
    --make-grm-bin \
    --out ukbio_10ksample_grm_chr${i} \
    --thread-num 16 > grm_${i}.log  \
done
```

Calculate the genetic variance explained using a general population gout prevalence of 2%
```{r, engine = 'bash',eval = FALSE}
for i in $(seq 1 22) 
do 
  gcta1.26.0 \
    --grm ukbio_10ksample_grm_chr${i} \
    --pheno phenos.txt  \
    --prevalence 0.02 \
    --out var_chr${i} \
    --reml \
    --chr ${i} \
    --thread-num 16 \
done

```

### Variant Annotation
snpEff


### Pathway Analysis


## Datasets

### 1000 Genomes Project Phase 3


@misc{1kgp_ftp,
keywords = {1KGP},
mendeley-groups = {PPARGC1Apaper,Datasets},
mendeley-tags = {1KGP},
title = {{1000 Genomes Project FTP}},
url = {ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/},
urldate = {2015-08-18}
}


### Axiom

- about
- population

| | obs missing (n) | Māori AXIOM
---|---|---
Number | - | 71 
Sex (% male) | - | 64.7
Gout cases (%) | - | 49.3
Mean BMI (range) | 1 | 34.5 (22.1 to 56.1)
Diabetes (%) | 3 | 21.1
Mean grandparent ancestry (range) | - | 0.916 (0.625 to 1.00)

Mean grandparent ancestry is the proportion of grandparents self-reported as belonging to the ethnicity group BMI body mass index


- markers
- density


### Omni


| | obs missing (n) | Samoan OMNIexpress
---|---|---
Number | - | 96
Sex (% male) | - | 74
Gout cases (%) | - | 50.0
Mean BMI (range) | - | 36.1 (20.5 to 93.4)
Diabetes (%) | - | 12.5
Mean grandparent ancestry (range) | - | 0.997 (0.875 to 1.00)

Mean grandparent ancestry is the proportion of grandparents self-reported as belonging to the ethnicity group BMI body mass index


- about
- population
- markers
- density

### CoreExome

The Infinium CoreExome-24 Beadchip is a SNP geneotyping platform available from Illumina and comprises a core set of 551,839 markers with an additional 100,000 custom markers able to be added. The markers are on average 5.26 kb apart.

Marker Categories | Number of Markers
---|---
In RefSeq Genes | 374,258 (399,009 within 10 kb)
In RefSeq Exons | 269,186
In RefSeq Promoter Regions | 13,990
In ADME Genes | 11,276 (12,455 within 10 kb)
In ADME Exons | 9,314
MHC (Extended MHC) | 3,940 (5,628)
Overlap with Genes in COSMIC | 365,135
Overlap with Genes in Gene Ontology | 96,331
Nonsense Markers | 14,994
Missense Markers | 219,162
Synonymous Markers | 14,747
Silent Markers | 8,850
Mitochondrial Markers |351
Indels | 12,436

mean call rate
mean missing
number of snps passing qc
number of snps used for selection


(http://www.illumina.com/content/dam/illumina-marketing/documents/products/datasheets/datasheet_human_core_exome_beadchip.pdf, accessed 16 Jan 2017)


Clinical Information


\newpage
\blandscape
```{r, echo = FALSE, results = 'asis', message = FALSE, warning= FALSE}
 knitr::kable(ce_clinical_table(), longtable = TRUE, format = 'pandoc', booktabs = TRUE , caption = 'Clinical information for each core Exome population')
```



\elandscape
\newpage

```{r, echo = FALSE, results = 'asis', message = FALSE, warning=FALSE}
knitr::kable(ce_populations_table(), longtable = TRUE, format = 'pandoc', booktabs = TRUE , caption = 'Clinical information for New Zealand Populations in selection analysis')
```


- about
- population
- markers
- density

### Polynesian Sequences

n | ~90
mean coverage |  Y x


### Reference datasets

\newpage
\blandscape

name | URL | date downloaded 
--- | --- | ---
1000 genomes phased haplotypes | https://mathgen.stats.ox.ac.uk/impute/1000GP_Phase3.html | 
hs37d5 | ftp://ftp.1000genomes.ebi.ac.uk:21/vol1/ftp/technical/reference/phase2_reference_assembly_sequence/hs37d5.fa.gz | 5 Aug 2014
Ancestral allele | ftp://ftp.ensembl.org/pub/release-66/fasta/ancestral_alleles/homo_sapiens_ancestor_GRCh37_e66.tar.bz | 6 Aug 2014 

\elandscape
\newpage